---
title: "Assign 3"
author: "Esther"
date: "20/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages & Data

```{r load packages, echo=FALSE}

library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)

```

```{r load data}

Hydrants <- st_read("https://data.cambridgema.gov/api/geospatial/szvy-s7ga?method=export&format=KML", 
                    quiet=TRUE)

MemorialPoles <- st_read("https://data.cambridgema.gov/api/geospatial/j2pj-hhcb?method=export&format=KML", 
                         quiet=TRUE)

Playground <- st_read("https://data.cambridgema.gov/api/geospatial/kmiz-sbu9?method=export&format=KML",
                      quiet=TRUE)

FStation <- st_read("https://data.cambridgema.gov/api/geospatial/q3m9-gvcq?method=export&format=KML", 
                    quiet = TRUE)

Neighborhood <- st_read("https://data.cambridgema.gov/api/geospatial/k3pi-9823?method=export&format=KML", 
                         quiet=TRUE)

OpenS <- st_read("https://data.cambridgema.gov/api/geospatial/88i7-ga9f?method=export&format=KML",
                 quiet=TRUE)



```


## Transform the Data

```{r}

MAState <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

Hydrants <- Hydrants %>%
  st_transform(MAState)

MemorialPoles <- MemorialPoles %>%
  st_transform(MAState)

FStation <- FStation %>%
  st_transform(MAState)

Playground <- Playground %>%
  st_transform(MAState)

OpenS <- OpenS %>%
  st_transform(MAState)

Neighborhood <- Neighborhood %>%
  st_transform(MAState)

```


## Map 1 - Hydrants within 350m of a Fire Station 

The number and proportion of A points within a specified distance of B points.
A Point-layer : Hydrants
B Point-layer : Fire Station

```{r}

FStation_buffer <- st_buffer(FStation, dist = 350) %>%
  st_union()

Hy_Fire <- Hydrants[FStation_buffer,]

Hydrants <- Hydrants %>%
  st_join(Hy_Fire) %>%
  mutate(by_FStation = !is.na(Name.y))

n_FStation_Hy <- sum(Hydrants$by_FStation)

n_Hydrants <- length(Hydrants$by_FStation)
pct_FStation_Hy <- n_FStation_Hy / n_Hydrants

left_side  <- st_bbox(Neighborhood)$xmin
top_side <- st_bbox(Neighborhood)$ymax

ggplot(Neighborhood) +
  geom_sf(fill = "papayawhip", color = "peachpuff4", lty = 4) +
  geom_sf(data = Hydrants, size = 0.2,
          aes(color = by_FStation)) +
  scale_color_manual(values = c("darksalmon", "darkred"),
          name = "Cambridge Hydrants\nby distance to a Fire Station", 
          labels = c("No Fire Station within 350 m",
                     "Fire Station within 30 m")) +
  geom_sf(data = FStation_buffer, fill = NA, color = "darkred", lty = 4) +
  geom_sf(data = FStation, color = "brown4", shape = 18， size = 3, alpha = 0.7) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_Hydrants, big.mark = ","),
                         " Hydrants in Cambridge,", 
                         prettyNum(n_FStation_Hy),
                         " (", 
                         prettyNum(100*pct_FStation_Hy, digits = 0),
                         "%) are within 350 meters of a parking meter.",
                         sep = ""),
           hjust = 0, vjust = 0.5, size = 2.5) +
  theme_map() +
  theme(panel.background = element_rect(fill = "floralwhite"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```

## Map 2 - Average number of Hydrants in each Neighborhood

The average number of A points in each C polygon
A Point : Hydrants
C Polygon : Neighborhood

```{r}

Neighborhood <- Neighborhood %>%
  mutate(num_Hy = lengths(st_covers(Neighborhood, Hydrants)))

ggplot(Neighborhood) +
  geom_sf(color = NA, 
          aes(fill = num_Hy)) +
  scale_fill_viridis_c(name = "Cambridge neighborhoods\nby number of Hydrants",
                       breaks = breaks <- seq(0, 250, by = 50),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "hydrants")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = "right",
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```

## Map 3 - Distance of Hydrants to Fire Station

The average (Euclidean) distance between A points and their nearest respective B points.
A Point - Hydrants
B Point - Fire Station

```{r}

right_side <- st_bbox(Hydrants)$xmax
left_side  <- st_bbox(Hydrants)$xmin
top_side <- st_bbox(Hydrants)$ymax
bottom_side <- st_bbox(Hydrants)$ymin

FStation <- FStation %>%
  mutate(Hy_dist = st_nn(FStation, Hydrants, 
                           returnDist = TRUE)$dist) %>%
  mutate(Hy_dist = as.numeric(Hy_dist))

avg_Hy_dist <- mean(FStation$Hy_dist)

ggplot(Neighborhood) +
  geom_sf(fill = "papayawhip", color = "peachpuff4", lty = 4) +
  
  geom_sf(data = FStation, size = 5, shape = 18, aes(color = Hy_dist)) +
  scale_color_viridis_c(name = 
                          "Cambridge Hydrants\nby distance to a Fire Station") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side + 300, 
           y = top_side, 
           label = paste("On average, a Cambridge Hydrant is ", 
                         prettyNum(avg_Hy_dist, digits = 0),
                         " meters from a Fire Station.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "floralwhite"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```

## Map 4 - Neighborhood with Open Space
The number and proportion of C polygons that overlap with D polygons.
C Polygon - Neighborhood
D Polygon - Open Space

```{r}

Neighborhood <- Neighborhood %>%
  mutate(num_OpenS = lengths(st_overlaps(Neighborhood, OpenS))) %>%
  mutate(has_OpenS = num_OpenS > 0)

n_OpenS_Neighborhood <- sum(Neighborhood$has_OpenS)

n_Neighborhood <- length(Neighborhood$num_OpenS)
pct_Neigh_OpenS <- n_OpenS_Neighborhood / n_Neighborhood

left_side  <- st_bbox(Neighborhood)$xmin
top_side <- st_bbox(Neighborhood)$ymax

ggplot(Neighborhood) +
  geom_sf(fill = "papayawhip", color = "white", lty = 4) +
  geom_sf(data = Neighborhood,
          aes(fill = has_OpenS)) +
  scale_fill_manual(values = c("cornsilk", "seagreen4"),
          name = "Cambridge Neighborhoods\nby presence of Open Space", 
          labels = c("Neighborhood without\nan overlapping body of Open Space",
                     "Neighborhood with an\noverlapping body of Open Space")) +
  geom_sf(data = OpenS, fill = "seagreen1", color = NA) + 
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste(n_OpenS_Neighborhood ,
                          "(", 
                         prettyNum(100*pct_Neigh_OpenS, digits = 0),
                         "%) of Cambridge's", 
                         length(Neighborhood$Name),
                         "neighborhoods contain or overlap with", 
                         "a body of Open Space."),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "floralwhite"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))

```









